Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-08-31 paper.tex
Edits

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Hierarchical Model for Expert Beliefs}

@@

We use this procedure to generate the hidden processes for $K$ questions with $T_k = 100$ for $k = 1, \dots, K$. Each question involves 50 experts allocated evenly among the five expertise groups. The expert logit-probabilities $Y_{t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$. Similarly to the real-world data setting discussed in Section \ref{realData}, extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99).

The simulation is then performed over a three-dimensional grid of parameter values:
\begin{eqnarray*}
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
\sigma^2 &\in& \{0.5, 1.0, 1.5, 2.0, 2.5\}\\
K &\in& \{20, 35, 50, 65, 80\},
\end{eqnarray*}
where $\beta$ generates the bias vector by  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$. 
The simulation runs a total of 40 times at every grid point. Each run records the accuracy to which the approaches are able to estimate the hidden process and bias vector. The first is measured in terms of the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} (\hat{p}_{t,k} - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, while the latter is measured in terms of the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.

@@

Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy. Even though some relative differences are visible, the loss values are very small in absolute terms; hence suggesting good overall accuracy to capture the hidden process. Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger. The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities and hence to more frequent truncation. This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions.

Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large. This behavior can be again considered as being caused  to a large extent by the truncation of the extreme expert logit-probabilities. Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.

@@

\subsection{Summary of Findings}
The model was applied to an unusually large dataset on expert probability forecasts. The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample settings. This has direct implications  on predictive power and model interpretability. Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability. In particular, the model was deemed a well-balanced compromise that avoids over-fitting without begin overly conservative. Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise. All the groups were found to be under-confident. The under-confidence, however, decreased as the level of self-reported expertise increased. This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}). Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific parameters. These quantities have clear interpretations and can be combined in many interesting ways to explore a range of hypotheses about different types of questions and expert behavior.

>>>>

\begin{document}
\begin{frontmatter}
% "Title of the paper"
\title{Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs}
\runtitle{Dynamic Hierarchical Modeling of Sparse Expert Beliefs}

@@

We use this procedure to generate the hidden processes for $K$ questions with $T_k = 100$ for $k = 1, \dots, K$. Each question involves 50 experts allocated evenly among the five expertise groups. The expert logit-probabilities $Y_{i,t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$. Any extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99). This is akin to the real-world data setting, where extreme probabilities must be truncated (see Section \ref{realData}). Therefore, to understand the effect that truncation has on the real-world data results, it should be first analyzed and understood under the controlled environment of synthetic data.

Throughout the evaluation data is generated under different choices of parameter values. More specifically, the evaluation procedure iterates over a three-dimensional grid of parameter values
\begin{eqnarray*}
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
\sigma^2 &\in& \{0.5, 1.0, 1.5, 2.0, 2.5\}\\
K &\in& \{20, 35, 50, 65, 80\},
\end{eqnarray*}
where $\beta$ generates the bias vector  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$. These values were considered to be realistic based on our experience with the real-world probability forecasting data. Each triple of parameter values,  i.e. a grid point, is used  40 times to generate a synthetic dataset. This dataset is then used to assess how accurately the approaches are able to estimate the hidden process and bias vector. The accuracy to estimate the hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy to capture the bias vector is quantified by the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.

@@

Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy. Even though some relative differences are visible, the loss values are very small in absolute terms; hence suggesting good overall accuracy to capture the hidden process. Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger. The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities and hence to more frequent truncation. This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions. Increasing the number of questions improves the estimation accuracy but only very little.

Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large. This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities. Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.

@@

\subsection{Summary of Findings}
The model was applied to an unusually large dataset on expert probability forecasts. The estimated crowd belief was found to be sharp and well-calibrated under both in- and out-of-sample settings. This has direct implications  on predictive power and model interpretability. Firstly, the model was shown to outperform other probability aggregators in terms of forecasting ability. In particular, the model was deemed a well-balanced compromise that avoids over-fitting without being overly conservative. Secondly, the crowd belief was used as the no-bias reference point to study the bias among groups of experts with different levels of self-reported expertise. All the groups were found to be under-confident. The under-confidence, however, decreased as the level of self-reported expertise increased. This result is about groups of experts and hence does not conflict with the well-known result of the individuals being over-confident (see, e.g.,  \citet{lichtenstein1977calibration, morgan1992uncertainty, bier2004implications}). Besides making predictions or studying group-level bias, the model can be used to generate estimates of many problem-specific parameters. These quantities have clear interpretations and can be combined in many interesting ways to explore a range of hypotheses about different types of questions and expert behavior.
